---
title: "RT analysis"
author: "Rebecca Lowdon"
date: "June 6, 2016"
output: html_document
---

Replication timing overlap analysis

# RT partitions
* RT partitions done via python script `RT_partitions.py`

```{r}
setwd("~/cancer_annotation/RT/v77/")

cmd01 = "python RT_partitions.py RT_CyT49_Liver_D16_All.txt > liver_D16_all.partitions.hg19.txt"
cmd02 = "python RT_partitions.py RT_CyT49_Liver_D16_All-2.txt > liver_D16_all-2.partitions.hg19.txt"

if( !(file.exists("liver_D16_all.partitions.hg19.txt"))){
  system( cmd01 )
}

## NTS 6/6: cmd02 returns on ly 169 lines for 2nd liver sample data.
if( !(file.exists("liver_D16_all-2.partitions.hg19.txt"))){
  system( cmd02 )
}
```

* Read in values and clean data.

```{r, echo=FALSE}
library(dplyr)
library(GenomicRanges)
library(GenomicFeatures)

rt1 <- read.table("liver_D16_all.partitions.hg19.txt",sep = "\t",header = F)
colnames(rt1) <- c("chr","start","end","direction")
  glimpse(rt1)
  
rt1_auto <- filter(rt1, chr!="chrY", chr!="chrX") #, chr!="chr15_KI270727v1_random")
  unique(rt1_auto$chr)
```

* Get num bp in each RT partition:

```{r}
liv1_stat <- mutate(rt1_auto, w=end-start) %>% group_by(., direction) %>% summarise(., sum(w))
  head(liv1_stat)
  tot <- 1478862868 + 1201145767 # for liver rep 1 #1420507557 + 1261494476 # for liver rep2 # 
  early <- 1201145767 / tot   ## 45% of dataset is early-replicating
  late <- 1478862868 / tot   ## 55% of dataset is in late-replicating

liv1_stat$bkgrd[liv1_stat$direction=="-1"] <- late
liv1_stat$bkgrd[liv1_stat$direction=="1"] <- early

liv1_stat
tot
```

* Plot distribution of bin lengths, by signal direction:

```{r}
d <- mutate(rt1_auto, w=end-start)
d$direction <- factor(d$direction)
  glimpse(d)
  
ggplot(d, aes(w, fill=direction)) + geom_histogram(position = position_dodge(), bins=100) 

png("liver1_RT_domains_boxpl.png")
ggplot(d, aes(x=direction,y=w)) + geom_boxplot() + scale_y_log10() + geom_jitter(alpha=0.4,width = 0.8) + theme_bw(base_size = 24) + ylab("Domain width") + xlab("RT domains")
dev.off();

group_by(d, direction) %>% summarise(., min(w), median(w), mean(w), sd(w), max(w))

```


* Make into a GRanges object for analysis in context of SNVs.

```{r}
liv1_auto_gr <- makeGRangesFromDataFrame(rt1_auto,keep.extra.columns = T)
  head(liv1_auto_gr)
  table(elementMetadata(liv1_auto_gr))
```


# Compare with filtered or removed SNVs

```{r}
df <- read.table(file = "~/cancer_annotation/COSMIC/WGSv77/liver_remove_WGS_hg19_v77_975.bed",header=F,sep="\t")
df <- read.table(file = "~/cancer_annotation/COSMIC/WGSv77/liver_keep_WGS_hg19_v77_975.bed",header=F,sep="\t")

colnames(df) <- c("chr","start","end","cosmicID","sampleID","ExSeq","FATHMM_score")
df <- filter(df, chr!="chrY", chr!="chrX")
  head(df)

SNVs <- makeGRangesFromDataFrame(df, keep.extra.columns = T, ignore.strand = T)
SNVs <- dropSeqlevels(SNVs, c("chr23","chr24") )
  seqlevels(SNVs)
  head(SNVs)
  length(SNVs)  # v77 hg19 kept: 2854 ; removed: 3872 (autosomes only!)
```

* Use overlap function to get # of SNVs in each `liv1_auto_gr` region.

```{r}
SNVs_by_RT <- as.data.frame(mergeByOverlaps(SNVs, liv1_auto_gr))
SNVs_by_RT <- dplyr::select(SNVs_by_RT, -cosmicID, -sampleID, -ExSeq, -FATHMM_score, -direction)
  glimpse(SNVs_by_RT)
  dim(SNVs_by_RT)
  length(unique(SNVs_by_RT$SNVs.cosmicID))

table(SNVs_by_RT$liv1_auto_gr.direction)
head(SNVs_by_RT)
```

* At this point, enter values into Chi-squared test in Excel.

```{r}
liv1_stat$snv_counts[liv1_stat$direction=="1"] <- 
length(SNVs_by_RT$SNVs.cosmicID[SNVs_by_RT$liv1_auto_gr.direction=="1"])/dim(SNVs_by_RT)[1]

liv1_stat$snv_counts[liv1_stat$direction=="-1"] <- 
length(SNVs_by_RT$SNVs.cosmicID[SNVs_by_RT$liv1_auto_gr.direction=="-1"])/dim(SNVs_by_RT)[1]

liv1_stat

mat <- cbind(liv1_stat$snv_counts,liv1_stat$bkgrd)
mat
fisher.test(mat,alternative = "g")

```


## Based on overlap done on command line
* using bash script

```{r}
d <- read.table("liver_keep_WGS_hg38.bed_vs_liver_R2",header = F,sep = "\t")
d <- read.table("liver_keep_WGS_hg19_v77_975__vs__liver_D16_partions",header = F)
colnames(d) <-c("chr","start","end","direction","count_snvs")
d$direction <- factor(d$direction)
glimpse(d)

d1 <- group_by(d, direction) %>% summarise(., sum(count_snvs))
d2 <- group_by(d, direction) %>% mutate(., w=end-start) %>% summarise(., sum(w))

d_stat <- cbind(d1,d2[,2])
d_stat
```

* Removed values

```{r}
r <- read.table("liver_remove_WGS_hg38.bed_vs_liver_R2",header = F,sep = "\t")
colnames(r) <-c("chr","start","end","direction","count_snvs")
r$direction <- factor(r$direction)
glimpse(r)

r1 <- group_by(r, direction) %>% summarise(., sum(count_snvs))
r2 <- group_by(r, direction) %>% mutate(., w=end-start) %>% summarise(., sum(w))

r_stat <- cbind(r1,r2[,2])

r_stat

```

* Then use Chi-squared test (e.g. in Excel: http://www.real-statistics.com/chi-square-and-f-distributions/independence-testing/)
